### 2015-04-14 abstract2.tex
chapter 2 little modify

Artificial intelligence is a very important subject in computer science. The knowledge representation and reasoning in artificial intelligence aims at completing and reflecting human intelligence in computer through the understanding of intelligence and cognitive nature. John McCarthy, one of the founders of artificial intelligence, has begun to promote the development of knowledge representation in the early 1950s. He first proposed the use of symbolic reasoning form to depict the cognitive and reasoning of knowledge.

Knowledge representation is through monotonic logic for reasoning at its early age. However, people came to realize that human cognition of common sense and reasoning is non-monotonic soon. In this case, non-monotonic logic became the mainstream tool of knowledge representation from then on. In this thesis, the answer set programming (ASP) which we studied is an import content of non-monotonic logic. With many years development of answer set programming problems, its theory and solver has been relatively mature. Nevertheless, the computational complexity of normal logic program (NLP) and disjunctive logic program (DLP) in ASP program under stab le model semantic are $\mathbf{NP}\text{-}complete$ and $\Pi^P_2\text{-}complete$ respectively after strict proven. Therefore, the main development bottleneck problem in answer set program is solver efficiency.

In the process of ASP solver speeding up development, Lifschitz and Turner proposed the concept of splitting set and program splitting in 1994. Moreover, they provided a method to divide a logic program into two parts which was named as bottom and top, and showed that the task of computing the answer sets of the program can be converted into the tasks of computing the answer sets of these parts. The concept of splitting set and program splitting brought new ideas to speed up solving answer sets of ASP program. In the subsequent period, splitting set and program splitting have been promoted and recognized continuously. However, the notion of splitting set which proposed by Lifschitz and Turner was based on strong conditions. Because of this, the empty set and the set of all atoms are the only two splitting sets for many ASP programs in the actual situation. These two splitting sets made no sense to speed up the answer sets solving in ASP programs, because these splitting sets cannot divide programs by the splitting methods.

Based on this background, in this thesis, I carried on the deep discussion and research about Lifschitz and Turner’s splitting set and program splitting theorem and extend them. Along this train of thought, I introduce the concept of splitting set to program simplification. The main achievements obtained in this thesis are shown as following details.

Firstly, I extend Lifschitz and Turner’s splitting set and program splitting theorem to allow the program to be split by an arbitrary set of atoms and introduce a new program splitting method. As the splitting set can be an arbitrary set of atoms, the applicability of program splitting can be extended greatly. Besides, this thesis figure out the main performance bottlenecks through analyzing properties of the new splitting method. At the same time, the data coming from experiment in this thesis support the fact that using arbitrary set of atoms as splitting set and the new splitting method to divide ASP program and solve its answer sets is quicker than solve directly. More precisely, it is two to three times faster than the original according the experiment.

Secondly, during analyzing how the new splitting set effect on the performance of new splitting method, I found out that if atoms in the splitting set are satisfied by every answer set of the program, it could release the computational complexity of answer set solving. According to this, this thesis extends the usage of splitting set to program simplification. In the same words, we can use consequence of the ASP programs to simplify themselves. The consequence of an ASP program is a set of literals that are satisfied by every answer set of the program. I took only positive literals to make them as atoms. The purpose of program simplification is still to speed up solving answer sets of ASP program.

This thesis proposes the new splitting set and new splitting method to make contribution to speed up solving answer sets of ASP program which brings a substantial progress, and extends the concept of splitting set to practical application such as program simplification. All these make important improvement to solving ASP program.

>>>>

Artificial intelligence is a very important subject in computer science. The knowledge representation and reasoning in artificial intelligence aims at completing and reflecting human intelligence in computer through the understanding of intelligence and cognitive nature. Non-monotonic logic is the mainstream tool of knowledge representation and reasoning. In this dissertation, the answer set programming (ASP) which we studied is an import content of non-monotonic logic. With many years development of answer set programming problems, its theory and solver has been relatively mature. However, the main development focus in answer set program is still solver efficiency.

In the process of ASP solver speeding up development, Lifschitz and Turner proposed the concept of splitting set and program splitting in 1994. Moreover, they provided a method to divide a logic program into two parts which was named as bottom and top, and showed that the task of computing the answer sets of the program can be converted into the tasks of computing the answer sets of these parts. The concept of splitting set and program splitting brought new ideas to speed up solving answer sets of ASP program. In the subsequent period, splitting set and program splitting have been promoted and recognized continuously. However, the notion of splitting set which proposed by Lifschitz and Turner was based on tough conditions. Because of this, the empty set and the set of all atoms are the only two splitting sets for many ASP programs in the actual situation. These two splitting sets made no sense to speed up the answer sets solving in ASP programs, because these splitting sets cannot divide programs by the splitting methods.

In this dissertation, I will discuss and research about Lifschitz and Turner’s splitting set and program splitting theorem. The main achievements obtained in this dissertation are shown as following details.

Firstly, I extend Lifschitz and Turner’s splitting set and program splitting theorem to allow the program to be split by an arbitrary set of atoms and introduce a new program splitting method for NLP. After this, I extend this result to DLP and propose the strong splitting method. As the splitting set can be an arbitrary set of atoms, the applicability of program splitting can be extended greatly.

Secondly, this dissertation figures out the main performance bottlenecks 
thr-ough analyzing properties of the new splitting method and puts forward the improved scheme. Moreover, the data coming from experiment in this dissertation support the fact that using arbitrary set of atoms as splitting set and the new splitting method to divide ASP program and solve its answer sets is quicker than solve directly. More precisely, it is two to three times faster than the original according the experiment.

Thirdly, during analyzing how the new splitting set effect on the performance of new splitting method, I found out that if atoms in the splitting set are satisfied by every answer set of the program, it could release the computational complexity of answer set solving. According to this, this dissertation extends the usage of splitting set to program simplification. In the same words, we can use consequence of the ASP programs to simplify themselves. The consequence of an ASP program is a set of literals that are satisfied by every answer set of the program. I take only positive literals to make them as atoms. The purpose of program simplification is still to speed up solving answer sets of ASP program.

This dissertation proposes the new splitting set and new splitting method to make contribution to speed up solving answer sets of ASP program which brings a substantial progress, and extends the concept of splitting set to practical application such as program simplification. All these bring new ideas to improve solving ASP program.

%Artificial intelligence is a very important subject in computer science. The knowledge representation and reasoning in artificial intelligence aims at completing and reflecting human intelligence in computer through the understanding of intelligence and cognitive nature. John McCarthy, one of the founders of artificial intelligence, has begun to promote the development of knowledge representation in the early 1950s. He first proposed the use of symbolic reasoning form to depict the cognitive and reasoning of knowledge.
%
%Knowledge representation is through monotonic logic for reasoning at its early age. However, people came to realize that human cognition of common sense and reasoning is non-monotonic soon. In this case, non-monotonic logic became the mainstream tool of knowledge representation from then on. In this thesis, the answer set programming (ASP) which we studied is an import content of non-monotonic logic. With many years development of answer set programming problems, its theory and solver has been relatively mature. Nevertheless, the computational complexity of normal logic program (NLP) and disjunctive logic program (DLP) in ASP program under stab le model semantic are $\mathbf{NP}\text{-}complete$ and $\Pi^P_2\text{-}complete$ respectively after strict proven. Therefore, the main development bottleneck problem in answer set program is solver efficiency.
%
%In the process of ASP solver speeding up development, Lifschitz and Turner proposed the concept of splitting set and program splitting in 1994. Moreover, they provided a method to divide a logic program into two parts which was named as bottom and top, and showed that the task of computing the answer sets of the program can be converted into the tasks of computing the answer sets of these parts. The concept of splitting set and program splitting brought new ideas to speed up solving answer sets of ASP program. In the subsequent period, splitting set and program splitting have been promoted and recognized continuously. However, the notion of splitting set which proposed by Lifschitz and Turner was based on strong conditions. Because of this, the empty set and the set of all atoms are the only two splitting sets for many ASP programs in the actual situation. These two splitting sets made no sense to speed up the answer sets solving in ASP programs, because these splitting sets cannot divide programs by the splitting methods. 
%
%Based on this background, in this thesis, I carried on the deep discussion and research about Lifschitz and Turner’s splitting set and program splitting theorem and extend them. Along this train of thought, I introduce the concept of splitting set to program simplification. The main achievements obtained in this thesis are shown as following details. 
%
%Firstly, I extend Lifschitz and Turner’s splitting set and program splitting theorem to allow the program to be split by an arbitrary set of atoms and introduce a new program splitting method. As the splitting set can be an arbitrary set of atoms, the applicability of program splitting can be extended greatly. Besides, this thesis figure out the main performance bottlenecks through analyzing properties of the new splitting method. At the same time, the data coming from experiment in this thesis support the fact that using arbitrary set of atoms as splitting set and the new splitting method to divide ASP program and solve its answer sets is quicker than solve directly. More precisely, it is two to three times faster than the original according the experiment.
%
%Secondly, during analyzing how the new splitting set effect on the performance of new splitting method, I found out that if atoms in the splitting set are satisfied by every answer set of the program, it could release the computational complexity of answer set solving. According to this, this thesis extends the usage of splitting set to program simplification. In the same words, we can use consequence of the ASP programs to simplify themselves. The consequence of an ASP program is a set of literals that are satisfied by every answer set of the program. I took only positive literals to make them as atoms. The purpose of program simplification is still to speed up solving answer sets of ASP program.
%
%This thesis proposes the new splitting set and new splitting method to make contribution to speed up solving answer sets of ASP program which brings a substantial progress, and extends the concept of splitting set to practical application such as program simplification. All these make important improvement to solving ASP program.

### 2015-04-14 chap2.tex
chapter 2 little modify

其中$r$代表形式(\ref{asp_rule})中的整个规则，
$head(r)$称为规则$r$的头部，$body(r)$称为规则$r$的体部。
具体有$head(r)$为$a_1 \lor a_2 \lor~...~\lor~a_k$，
$body(r)$为$a_{k+1},~...~,~a_m,~not~a_{m+1},~...~, \\ not~a_n$。
其中ASP逻辑程序的头部中的原子连接关系只为析取，体部中的原子连接关系只为合取。
更进一步的划分有$body(r)~=~body^+ (r)~\land~body^-(r)$，其中
$body^+(r)$为$a_{k+1},~...~,~a_m$，即原规则中不带$not$的文字集合，
$body^-(r)$为$not~a_{m+1},~ \\ ...~,~not~a_n$，即原规则中带有$not$的文字集合。
不失一般性地，上述定义的命题公式可以被用作原子集（$a~set~of~atoms$）进行讨论，
并引入以下两个常用原子集：
\begin{definition}\upcite{lifschitz2002answer}
	给定$r$为符合形式(\ref{asp_rule})的一个规则，$P$为一个ASP逻辑程序：	
	\begin{itemize}
		\item $Atoms(r) = head(r) \cup body^+(r) \cup body^-(r)$；
		\item $Atoms(P)~=\bigcup_{r \in P}Atoms(r)$。
	\end{itemize}
\end{definition}

\begin{example}
	把例子\ref{first_example_2}中的规则集合看作是一个逻辑程序$P$，对于(\ref{rule_5})则有：
	\begin{eqnarray}
	head(r)=\{s,q\}	\\
	body(r)=\{r,t,p\}	\\
	body^+ (r)=\{r,t\}	\\
	body^-(r)=\{p\}		\\
	Atoms(r)=\{s,q,r,t,p\}
	\end{eqnarray}

@@

Gelfond和Lifschitz(1988)提出了稳定模型语义(Stable Model Semantic)，
稳定模型语义解决了非单调推理无法解释失败即否定的问题，此外他们还给出了一个规约方法(G-L规约)，以化简一个ASP逻辑程序中的失败即否定。
\begin{definition}[G-L规约\upcite{gelfond1988stable}]
	给定原子集$S$和不含约束的ASP逻辑程序$P$，$P$基于$S$的G-L规约结果记为$P^S$。$P$通过以下两个化简规则得到$P^S$：
	\begin{itemize}
		\item 若一个规则的体部中有$not~p$，且$p \in S$，则删掉该规则；
		\item 对剩下的所有规则，删除体部中的$not~p$，$p$为$Atoms(P)$中任意一个原子。
	\end{itemize}
\end{definition}

@@

显然，通过G-L规约进行化简后得到的$P^S$是一个不包含任何失败即否定的ASP逻辑程序。这样的ASP逻辑程序只有一个唯一的极小模型，这个模型被称为稳定模型(Stable Model)，并记为$Cons(P^S)$\upcite{gelfond1988stable}。

@@

在引入具体的环及环公式定义前，需要先给出一个ASP逻辑程序的正依赖图(Positive Dependence Graph)及强连通分量(Strongly Conntected Component)的定义。
\begin{definition}[正依赖图(Positive Dependency Graph)\upcite{lin2004assat}]
	给定ASP逻辑程序$P$，以$P$中的原子作为顶点，规则作为构成边的依据，可以构造出一个有向连通图，称其为$P$的正依赖图，记为$G_P$。其中，当存在$P$中的一个规则形如$p \in head(r)$且$q \in body^+(r)$，则正依赖图中存在一条从原子$p$指向原子$q$的有向边。
\end{definition}

@@

\begin{example}
	给定逻辑程序$P$如下：
	\begin{eqnarray}
		a~\leftarrow~not~d.	\\
		d~\leftarrow~not~c.	\\
		a~\leftarrow~c,~d.	\\
		c~\leftarrow~a.
	\end{eqnarray}

@@

Lifschitz和Turner在给出分割集的概念后，也基于分割集提出了相应的程序分割(Program Splitting)方法。
程序分割方法具体就是根据定义找出分割集，通过分割集把原程序分割为$bottom$和$top$两部分，
并且证明原程序的回答集可以通过$bottom$和$top$这两部分的回答集所求得。
\begin{definition}[分割集(Splitting Set)\upcite{lifschitz1994splitting}]\label{splitting_set}
	给定一个ASP逻辑程序$P$，其分割集是一个原子集，记为$U$，$U$需要满足：对于任意的规则$r \in P$有$head(r) \cap U \neq \emptyset$蕴涵$Atoms(r)\subseteq U$。
\end{definition}

@@

基于分割集划分而成的$bottom$和$top$两部分的具体定义如下：
\begin{definition}[底部和顶部(bottom and top) \upcite{lifschitz1994splitting}] \label{bottom_top}
	给定一个ASP逻辑程序$P$，及其分割集$U$，标记$P$的底部为$b_U(P)$，顶部为$t_U(P)$，并定义为：
	\begin{eqnarray}
		b_U(P) = \{ r \in P~|~head(r) \cap U \neq \emptyset \}	\\
		t_U(P) = P~\backslash~b_U(P)
	\end{eqnarray}
\end{definition}

显然，ASP逻辑程序的$bottom$是把分割集$U$中的原子作为头部的规则都抽取出来，而$top$则是头部与分割集$U$无关的规则集合，且$\emptyset$和$Atoms(P)$是任意一个逻辑程序P的分割集。

为了通过$bottom$和$top$求解原程序的回答集，需要引入化简操作。
Lifschitz和\\Turner定义了操作$e_U(P,X)$来联合$bottom$的回答集和$top$求出另一部分的回答集。
\begin{definition}\label{e_U_P_X}
	给定ASP逻辑程序$P$，和原子集$X$、$U$，定义$e_U(P,X)$如下：
	\begin{itemize}
		\item 删除符合以下条件的规则$r$：$head(r) \cap X \neq \emptyset$且$body^+(r) \cap U \nsubseteq X$，或者$(body^-(r) \cap U) \cap X \neq \emptyset$；
		\item 在剩下的规则的体部中把所有形如$a$或$not~a$，其中$a \in U$的文字删掉。
	\end{itemize}
\end{definition}

@@

\begin{example}
	给定ASP逻辑程序$P$：
	\begin{eqnarray}
		a~\leftarrow~not~d.	\\
		d~\leftarrow~not~c.	\\
		a~\leftarrow~c,~d.	\\
		c~\leftarrow.
	\end{eqnarray}

>>>>

其中$r$代表形式(\ref{asp_rule})中的整个规则，
$head(r)$称为规则$r$的头部，$body(r)$称为规则$r$的体部。
具体有$head(r)$为$a_1 \lor a_2 \lor~...~\lor~a_k$，
$body(r)$为$a_{k+1},~...~,~a_m,~not~a_{m+1},~...~, \\ not~a_n$。
其中ASP逻辑程序的头部中的原子连接关系只为析取，体部中的原子连接关系只为合取。
更进一步的划分有$body(r)~=~body^+ (r)~\land~body^-(r)$，其中
$body^+(r)$为$a_{k+1},~...~,~a_m$，即原规则中不带$not$的文字集合，
$body^-(r)$为$not~a_{m+1},~ \\ ...~,~not~a_n$，即原规则中带有$not$的文字集合。
不失一般性地，上述定义的命题公式可以被用作原子集(a set of atoms)进行讨论，
并引入以下两个常用原子集：
\begin{definition}\upcite{lifschitz2002answer}
	给定$r$为符合形式(\ref{asp_rule})的一个规则，$P$为一个ASP逻辑程序：	
	\begin{itemize}
		\item $Atoms(r) = head(r) \cup body^+(r) \cup body^-(r)$；
		\item $Atoms(P)~=\bigcup_{r \in P}Atoms(r)$。
	\end{itemize}
\end{definition}

\begin{example}
	把例子\ref{first_example_2}中的规则集合看作是一个逻辑程序$P$，对于(\ref{rule_5})则有：
	\begin{eqnarray}
	&& head(r)=\{s,q\}	\\
	&& body(r)=\{r,t,p\}	\\
	&& body^+ (r)=\{r,t\}	\\
	&& body^-(r)=\{p\}		\\
	&& Atoms(r)=\{s,q,r,t,p\}
	\end{eqnarray}

@@

Gelfond和Lifschitz(1988)提出了稳定模型语义，
稳定模型语义解决了非单调推理无法解释失败即否定的问题，此外他们还给出了一个规约方法(G-L规约)，以化简一个ASP逻辑程序中的失败即否定。
\begin{definition}[G-L规约\upcite{gelfond1988stable}]
	给定原子集$S$和不含约束的ASP逻辑程序$P$，$P$基于$S$的G-L规约结果记为$P^S$。$P$通过以下两个化简规则得到$P^S$：
	\begin{itemize}
		\item 若一个规则的体部中有$not~p$，且$p \in S$，则删掉该规则；
		\item 对剩下的所有规则，删除体部中的$not~p$，$p$为$Atoms(P)$中任意一个原子。
	\end{itemize}
\end{definition}

@@

显然，通过G-L规约进行化简后得到的$P^S$是一个不包含任何失败即否定的ASP逻辑程序。这样的ASP逻辑程序只有一个唯一的极小模型，这个模型被称为稳定模型，并记为$Cons(P^S)$\upcite{gelfond1988stable}。

@@

在引入具体的环及环公式定义前，需要先给出一个ASP逻辑程序的正依赖图及强连通分量的定义。
\begin{definition}[正依赖图(Positive Dependency Graph)\upcite{lin2004assat}]
	给定ASP逻辑程序$P$，以$P$中的原子作为顶点，规则作为构成边的依据，可以构造出一个有向连通图，称其为$P$的正依赖图，记为$G_P$。其中，当存在$P$中的一个规则形如$p \in head(r)$且$q \in body^+(r)$，则正依赖图中存在一条从原子$p$指向原子$q$的有向边。
\end{definition}

@@

\begin{example}
	给定逻辑程序$P$如下：
	\begin{eqnarray}
		&& a~\leftarrow~not~d.	\\
		&& d~\leftarrow~not~c.	\\
		&& a~\leftarrow~c,~d.	\\
		&& c~\leftarrow~a.
	\end{eqnarray}

@@

Lifschitz和Turner在给出分割集的概念后，也基于分割集提出了相应的程序分割(Program Splitting)方法。
程序分割方法具体就是根据定义找出分割集，通过分割集把原程序分割为bottom和top两部分，
并且证明原程序的回答集可以通过bottom和top这两部分的回答集所求得。
\begin{definition}[分割集(Splitting Set)\upcite{lifschitz1994splitting}]\label{splitting_set}
	给定一个ASP逻辑程序$P$，其分割集是一个原子集，记为$U$，$U$需要满足：对于任意的规则$r \in P$有$head(r) \cap U \neq \emptyset$蕴涵$Atoms(r)\subseteq U$。
\end{definition}

@@

基于分割集划分而成的bottom和top两部分的具体定义如下：
\begin{definition}[底部和顶部(bottom and top) \upcite{lifschitz1994splitting}] \label{bottom_top}
	给定一个ASP逻辑程序$P$，及其分割集$U$，标记$P$的底部为$b_U(P)$，顶部为$t_U(P)$，并定义为：
	\begin{eqnarray}
		b_U(P) = \{ r \in P~|~head(r) \cap U \neq \emptyset \}	\\
		t_U(P) = P~\backslash~b_U(P)
	\end{eqnarray}
\end{definition}

显然，ASP逻辑程序的bottom是把分割集$U$中的原子作为头部的规则都抽取出来，而top则是头部与分割集$U$无关的规则集合，且$\emptyset$和$Atoms(P)$是任意一个逻辑程序P的分割集。

为了通过bottom和top求解原程序的回答集，需要引入化简操作。
Lifschitz和\\Turner定义了操作$e_U(P,X)$来联合bottom的回答集和top求出另一部分的回答集。
\begin{definition}\label{e_U_P_X}
	给定ASP逻辑程序$P$，和原子集$X$、$U$，定义$e_U(P,X)$如下：
	\begin{itemize}
		\item 删除符合以下条件的规则$r$：$head(r) \cap X \neq \emptyset$且$body^+(r) \cap U \nsubseteq X$，或者$(body^-(r) \cap U) \cap X \neq \emptyset$；
		\item 在剩下的规则的体部中把所有形如$a$或$not~a$，其中$a \in U$的文字删掉。
	\end{itemize}
\end{definition}

@@

\begin{example}
	给定ASP逻辑程序$P$：
	\begin{eqnarray}
		&& a~\leftarrow~not~d.	\\
		&& d~\leftarrow~not~c.	\\
		&& a~\leftarrow~c,~d.	\\
		&& c~\leftarrow.
	\end{eqnarray}
